---
title: "Alluvial Diagram with ggalluvial::CheatSheet"
output:
  html_document: default
  pdf_document: default
---


```{css, echo=FALSE}
h2,h3 {
  font-size: 32px;
  font-weight: bold;
  color: #FFA62F;
  border: 1px solid #FFA62F;
  background-color: #F0FFFF;
}


h1.title {
	color:#7F462C;
	font-family:Impact, Charcoal, sans-serif;
	font-size:40pt;
}

/*
.column1{
  float: left;
  width: 33%;
  text-align: left;
}
.column2{
  display: inline-block;
  width: 33%;
  text-align: left;
}
.column3{
  float: right;
  width: 33%;
  text-align: left;
}

code.hljs{
 font-size: 3px;
}
*/

```

<div class="column1">
## Introduction

**Alluvial Diagrams** are a type of **flow diagram** which can be used to visualize frequency distributions over time or frequency tables involving several categorical variables.

## Installation

```{r eval=F,message=F,warning=F}
install.packages("ggalluvial")
```



## Package Load


```{r message=F,warning=F}
library(ggalluvial)
library(patchwork)
```

## What're in the Package?

 - Geom/Stat related functions
   - geom_alluvium (GeomStratum) stat_alluvium (StatAlluvium)
   - geom_flow (GeomFlow) stat_flow (StatFlow)
   - geom_lode (GeomLode)
   - geom_stratum (GeomAlluvium) stat_stratum (StatStratum)
 - data transform functions
   - to_alluvia
   - to_alluvia_form
   - to_lodes
   - to_lodes_form
   - is_alluvia_form
   - is_alluvial
   - is_alluvial_alluvia
   - is_alluvial_lodes
   - is_lodes_form
 - datasets predefined
   - majors
   - vaccinations

</div>


<div class="column2">
## Data Structure And Data Used 

ggalluvial recognizes two formats of **“alluvial data”**:

 - Alluvia (wide) format
 
```{r eval=F}
is_alluvia_form(data, ..., axes = NULL, weight = NULL, logical = TRUE,silent = FALSE)
to_alluvia_form(data, key, value, id, distill = FALSE)
```
 
 - Lodes (long) format

```{r eval=F}
is_lodes_form(data, key, value, id, weight = NULL, site = NULL, logical = TRUE,silent = FALSE) 
to_lodes_form(data, ..., axes = NULL, key = "x", value = "stratum",id = "alluvium", diffuse = FALSE, discern = FALSE)
```


## Example
Using dataset of **Titanic** for illusitration.

```{r}
data(Titanic)
dat<-as.data.frame(Titanic)
```

**Titanic** is in array(tabular) form which stores multiple categorical dimensions such as

- Sex
- Age
- Class
- Survived

For convinience, base::data.frame() was used to convert such an **array** to an acceptable **data.frame**.


```{r eval=F}
is_alluvia_form(dat)
```
</div>

<div class="column3">

## Two Categorical Dimensions

Using <u>***Sex***</u> and <u>***Survived***</u> variables for illustration: 

```{r fig.width=9}
pp1<-ggplot(dat,aes(y = Freq, axis1 = Sex, axis2 = Survived)) +
        geom_alluvium(aes(fill=Sex),width = 1/12)+
        geom_stratum(width = 1/12, fill = "#6D7B8D", color = "#2B1B17") +
        geom_label(stat = "stratum", aes(label = after_stat(stratum))) +
        scale_x_discrete(limits = c("Sex", "Survived"), expand = c(.05, .05)) +
        theme_minimal()+
		ggtitle('Alluival flows filled with Sex') +labs(y='')
  
pp2<-ggplot(dat,aes(y = Freq, axis1 = Sex, axis2 = Survived)) +
        geom_alluvium(aes(fill=Survived),width = 1/12)+
        geom_stratum(width = 1/12, fill = "#6D7B8D", color = "#2B1B17") +
        geom_label(stat = "stratum", aes(label = after_stat(stratum))) +
        scale_x_discrete(limits = c("Sex", "Survived"), expand = c(.05, .05)) +
		theme_minimal()+
        ggtitle('Alluival flows filled with Survived') +labs(y='')


pp1+pp2
```  

Elements of above <span style='color:#254117'>**Alluvial Diagram**</span>:


 - **Axes**: multiple horizontally-distributed bars representing factor variables, such as  <u>***Sex***</u> and <u>***Survived***</u>

 - **Strata**: vertical divisions of these axes representing these variables' values such as <u>***Male/Female(Sex)***</u> and <u>***yes/no(Survived)***</u>

 - **Alluvial flows**: splines  connecting vertical subdivisions (**lodes**) within strata of adjacent axes representing subsets or amounts of observations that take the corresponding values of the corresponding variables. 


So propotion of strata at each axis and their connections can be visualizied from be Alluvial plot.

Alluvial flows can be shown according to **Aesthetic mappings**, in above figure, Sex and Survived were used to "fill" the **flow**.



## Third Categorical Dimension Added

For the **Titanic** data shown above in Alluvial Diagram, how do we introduce a  **3<sup>rd</sup>** variable(dimension)?

<u>***Age***</u> will be used as the **3<sup>rd</sup>** variable for illustration.

### **3<sup>rd</sup>** variable as **Aesthetic Mapping**

```{r fig.width=5}

ggplot(dat,aes(y = Freq, axis1 = Sex, axis2 = Survived)) +
        geom_alluvium(aes(fill=Age),width = 1/12)+
        geom_stratum(width = 1/12, fill = "#6D7B8D", color = "#2B1B17") +
        geom_label(stat = "stratum", aes(label = after_stat(stratum))) +
        scale_x_discrete(limits = c("Sex", "Survived"), expand = c(.05, .05)) +
        theme_minimal()+
		ggtitle('Alluival flows filled with Age') +labs(y='')
		
```

For **Sex** axis, the **lobes** were (from top to bottom):

- Sex(Male)-- Age(Child) --Survived(no)
- Sex(Male)-- Age(Adult) --Survived(no)
- Sex(Male)-- Age(Child) --Survived(yes)
- Sex(Male)-- Age(Adult) --Survived(yes)
- Sex(Female)-- Age(Child) --Survived(no)
- Sex(Female)-- Age(Adult) --Survived(no)
- Sex(Female)-- Age(Child) --Survived(yes)
- Sex(Female)-- Age(Adult) --Survived(yes)

### **3<sup>rd</sup>** Variable as Facets.


```{r fig.width=9}
ggplot(dat,aes(y = Freq, axis1 = Sex, axis2 = Survived)) +
        geom_alluvium(aes(fill=Sex),width = 1/12)+
        geom_stratum(width = 1/12, fill = "#6D7B8D", color = "#2B1B17") +
        geom_label(stat = "stratum", aes(label = after_stat(stratum))) +
        scale_x_discrete(limits = c("Sex", "Survived"), expand = c(.05, .05)) +
        theme_minimal()+
		ggtitle('Alluival flows filled with Sex') +labs(y='') +
		facet_wrap(~ Age, scales = "free_y") +
		coord_cartesian(clip = 'off')

```

### **3<sup>rd</sup>** Variable as **3<sup>rd</sup>** Axis.

```{r fig.width=6.4}

ggplot(dat,aes(y = Freq, axis1 = Sex, axis2 = Survived, axis3=Age)) +
        geom_alluvium(aes(fill=Age),width = 1/12)+
        geom_stratum(width = 1/12, fill = "#6D7B8D", color = "#2B1B17") +
        geom_text(stat = "stratum", aes(label = after_stat(stratum)),size=4) +
        scale_x_discrete(limits = c("Sex", "Survived","Age"), expand = c(.05, .05)) +
        theme_minimal()+
		ggtitle('Alluival flows filled with Age, 3 axes') +labs(y='') +
		coord_cartesian(clip = 'off')
		
```

## Reference

[ggalluvial:http://corybrunson.github.io/ggalluvial/](http://corybrunson.github.io/ggalluvial/)

</div>